---
output: 
  html_document:
    css: "css/cheat_style.css"
---

<div style="margin-top: -99px; color: #FFAA2D; margin-bottom: -10px; font-size: 30px;">
## R for Environmental Data
</div>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

<div class="clearbox" style="">
<div class="header"> GET STARTED </div>
<div class="fullbox" style="">
<div class="blue-header"> New Project </div>
`File` >`New Project` > `New Directory`


<div class="blue-header"> New R script</div>

`File` > `New File` > `R Script`

<div class="blue-header"> Where am I? </div>
```{r}
# Show current working dir
getwd()
```

__Files tab__ -> **More** -> _Go To Working Directory_


<div class="blue-header"> Install new packages </div>
```{r}
install.packages("readr")
library(readr)
```
</div></div>


<div class="clearbox">
<div class="header"> STORE VALUES </div>
<div class="linebox">
```{r}
# Use the left-arrow
miles <- 355

# Text goes in quotes
my_name <- "Yoda"

# Multiple values go in c()
my_pets <- c("Sparky", "Garfield", "Bumpky")

# Avoid numbers, spaces, & symbols: #,*,!
3*mice <- "bad_name"
```
</div></div>


<div class="clearbox" style="margin-left: 20px;">
<div class="header"> READ DATA </div>

<div class="linebox">
```{r}
library(readr)
my_data <- read_csv("txt_file.csv")
```
<i class="fa fa-file-text-o" aria-hidden="true" style="float: right;"></i><i class="fa fa-file-text" aria-hidden="true" style="float: right;"></i>

<p>- - - - - </p>

```{r}
library(readxl)
my_data <- read_excel("Excel_file.xlsx")
```
<i class="fa fa-file-text-o" aria-hidden="true" style="float: right;"></i><i class="fa fa-file-text" aria-hidden="true" style="float: right;"></i>
</div>


<div class="header"> EXPLORE </div>
<div class="linebox">
```{r}
nrow(my_data)
names()
summary()
library(dplyr)
glimpse()
```
</div>
</div>


<div class="clearbox" style="">
<div class="header"> CLEAN NAMES</div>
<div class="linebox">
```{r}
# Simplify all column names
library(janitor)
my_data <- clean_names(my_data)

# Assign new names
library(dplyr)
# New name goes on left: new_name = `curent name` 
my_data <- rename(my_data, ozone = `O3 ConCentraions ppb`)
```
</div></div>


```{r, echo=F}
<div class="box">
<div class="header"> Get R! </div>
RStudio is a free working environment for R. It helps visualize the data you're analyzing, keep projects organized and share your work.

<div class="fullbox" style="background-color: lightgrey;">
__Install R__
https://cran.r-project.org/bin/windows/base/

__Install RStudio__
https://www.rstudio.com/products/rstudio/download/

</div></div>
```


<div style="clear: both;"></div>

<div class="clearbox" style="">
<div class="header"> FILTER </div>
<div class="linebox">
```{r}
library(dplyr)
# Only keep rows with ozone greater than zero
filter(my_data, ozone > 0)
# Keep rows with site name Fond du Lac
filter(my_data, site_name == "Fond du Lac")
# Keep sites Fond du Lac or Pechanga
filter(my_data, site_name %in% c("Fond du Lac", "Pechanga"))
```
</div>


<div class="header"> HELP! </div>
<div class="linebox">
```{r}
# Function help
?read_csv
# Package help
??readr
# Search help
help.search("boxplot")
```
</div>
</div>


<div class="clearbox" style="">
<div class="header"> SUMMARIZE </div>
<div class="linebox">
```{r}
library(dplyr)
# Summarize entire table
summarize(my_data, avg_ozone = mean(ozone))
# Summary by site
group_by(my_data, site_id) %>%
   summarize(avg_ozone = mean(ozone))
```
</div>


<div class="header"> SHORTCUTS </div>
<div class="linebox">

- Run line
- Save script
- left arrow

Read data

- point click option

</div>

</div>

<div class="clearbox" style="width: 33%; margin-left: 20px;">
<div class="header"> PLOTS </div>
<div class="linebox" style="background-color: #FFAA2D;">
```{r}
library(ggplot2)

ggplot(my_data, aes(x = date, y = ozone)) +
  geom_point() +
  theme_minimal()
```

```{r}
# Add horiz line + title
ggplot(my_data, aes(x = date, y = ozone)) +
  geom_point() +
  geom_hline() +
  titles("Ozone 2017", "subtitle", caption = "reputable data source")
```

```{r}
ggplot(my_data) + geom_line()

ggplot(my_data) + facet_wrap()
```
</div></div>


<div class="clearbox" style="width: 16%; border-radius: 4px; margin-right: 0;">
<div class="header"> &nbsp; </div>
<div class="linebox" style="background-color: #FFAA2D;">
Cool plot 

<br>

Cool plot2

<br>

Magic magic
</div></div>


<div style="clear: both;"></div>


<div class="clearbox" style="width: 29.2%;">
<div class="fullbox">
<div class="header" style="margin-bottom: 6px;"> Instructors </div>

![](https://image.flaticon.com/icons/svg/194/194638.svg){width=34 style="float: left; margin-right: 12px;"} [Melinda.Ronca-Battista@nau.edu](Melinda.Ronca-Battista@nau.edu)

![](https://image.flaticon.com/icons/svg/194/194623.svg){width=34 style="float: left; margin-right: 12px;"} [Kristie.Ellickson@state.mn.us](kristie.ellickson@state.mn.us)

![](https://image.flaticon.com/icons/svg/194/194638.svg){width=34 style="float: left; margin-right: 12px;"} [VallenC@grandportage.com](Vallenc@grandportage.com)

![](https://image.flaticon.com/icons/svg/194/194623.svg){width=34 style="float: left; margin-right: 12px;"} [Dorian.Kvale@state.mn.us](dorian.kvale@state.mn.us)

![](https://image.flaticon.com/icons/svg/194/194623.svg){width=34 style="float: left; margin-right: 12px;"} [Jaime.Yazzie@nau.edu](Jaime.Yazzie@nau.edu)

</div></div>

<div class="clearbox" style="width: 19.2%;">
<div class="bluebox" style="width: 100%;">
<div class="header"> &nbsp; </div>

<p style="color: white; font-size: 24px; font-weight: 800;">
"THIS IS A PLAIN BLUE BOX WITH A PROFOUND STATEMENT"
</p>
<p class="reference">
&#8212;__A wise__,ividual
</p>
</div></div>

<div class="clearbox" style="margin-left: 20px; margin-top: -40px;">
<div class="header"> SAVE PLOTS </div>
<div class="linebox">
```{r}
library(ggsave)
# Save last plot
ggsave("2019-02-012_ozone_plot.png")
# Save an earlier plot that you named
ggsave(named_plot, "2019-02-012_ozone_plot.png")
```
</div></div>


<div class="clearbox" style="margin-top: -40px;">
<div class="header"> SAVE DATA </div>
<div class="linebox">
```{r}
library(readr)
# Save to CSV text file
write_csv(my_data, "my_cleaned_data.csv")
```
</div></div>




